---
title: "TE CLOSEST GENES"
author: "Joshua G. Medina-Feliciano"
date: "2024-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LOADING LIBRARY

```{r}

library(ggrepel)
library(patchwork)
library(dplyr)


# Read the TSV file into a data frame without header
genes_10k <- read.delim("01_DISTANCE_10K/10k_mapped2trxpt.tsv", header = FALSE)
colnames(genes_10k) <- c("genome_pred_id", "Trinity")
genes_5k <- read.delim("02_DISTANCE_5K/5k_mapped2trxpt.tsv", header = FALSE)
colnames(genes_5k) <- c("genome_pred_id", "Trinity")
genes_2k <- read.delim("03_DISTANCE_2K/2k_mapped2trxpt.tsv", header = FALSE)
colnames(genes_2k) <- c("genome_pred_id", "Trinity")
genes_adjacent <- read.delim("adjacentGenes_mapped2trxpt.tsv", header = FALSE)
colnames(genes_adjacent) <- c("genome_pred_id", "Trinity")

# Read the CSV files into data frames with headers
hpe12 <- read.csv("00_EXPRESSION/HPE_12.csv", header = TRUE, row.names = 1)
dpe1 <- read.csv("00_EXPRESSION/DPE_1.csv", header = TRUE, row.names = 1)
dpe3 <- read.csv("00_EXPRESSION/DPE_3.csv", header = TRUE, row.names = 1)
dpe7 <- read.csv("00_EXPRESSION/DPE_7.csv", header = TRUE, row.names = 1)
dpe14 <- read.csv("00_EXPRESSION/DPE_14.csv", header = TRUE, row.names = 1)
dpe14a <- read.csv("00_EXPRESSION/DPE_14_A.csv", header = TRUE, row.names = 1)
dpe14p <- read.csv("00_EXPRESSION/DPE_14_P.csv", header = TRUE, row.names = 1)
dpe21 <- read.csv("00_EXPRESSION/DPE_21.csv", header = TRUE)

```



```{r}

# Extract the IDs from the second column of the TSV file
trxpts_10k_ids <- genes_10k$Trinity

# Define a function to filter each CSV data frame with partial matching
filter_csv <- function(df, ids) {
  # Extract the part of the Trinity ID without the _i part
  df$Trinity_short <- sub("_i.*", "", df$Trinity)
  ids_short <- sub("_i.*", "", ids)
  
  # Perform the filtering with partial matching
  df_filtered <- df[df$Trinity_short %in% ids_short & df$padj < 0.05 & df$log2FoldChange >= 2, ]
  
  # Remove the temporary Trinity_short column
  df_filtered$Trinity_short <- NULL
  
  return(df_filtered)
}

# # Define a function to filter each CSV data frame
# filter_csv <- function(df, ids) {
#   df_filtered <- df[df$Trinity %in% ids & df$padj < 0.05 & df$log2FoldChange >= 2, ]
#   return(df_filtered)
# }

```




## 10K

```{r}

# Apply the filter function to each CSV data frame
hpe12_filtered <- filter_csv(hpe12, trxpts_10k_ids)
dpe1_filtered <- filter_csv(dpe1, trxpts_10k_ids)
dpe3_filtered <- filter_csv(dpe3, trxpts_10k_ids)
dpe7_filtered <- filter_csv(dpe7, trxpts_10k_ids)
dpe14_filtered <- filter_csv(dpe14, trxpts_10k_ids)
dpe14a_filtered <- filter_csv(dpe14a, trxpts_10k_ids)
dpe14p_filtered <- filter_csv(dpe14p, trxpts_10k_ids)
dpe21_filtered <- filter_csv(dpe21, trxpts_10k_ids)

#DROP NA ROWS

# Define a function to drop rows with NA in all columns
drop_na_rows <- function(df) {
  df[rowSums(is.na(df)) != ncol(df), ]
}

# Apply this function to each filtered data frame
hpe12_filtered <- drop_na_rows(hpe12_filtered)
dpe1_filtered <- drop_na_rows(dpe1_filtered)
dpe3_filtered <- drop_na_rows(dpe3_filtered)
dpe7_filtered <- drop_na_rows(dpe7_filtered)
dpe14_filtered <- drop_na_rows(dpe14_filtered)
dpe14a_filtered <- drop_na_rows(dpe14a_filtered)
dpe14p_filtered <- drop_na_rows(dpe14p_filtered)
dpe21_filtered <- drop_na_rows(dpe21_filtered)


```



```{r}


# List of data frame names as strings
df_10k <- c(
  'hpe12_filtered', 'dpe1_filtered', 'dpe3_filtered', 'dpe7_filtered',
  'dpe14_filtered', 'dpe14a_filtered', 'dpe14p_filtered', 'dpe21_filtered')

# Loop through each data frame name and perform the join operation
for (name in df_10k) {
  # Get the data frame by name
  df <- get(name)
  
  # Perform the inner join with genes_10k
  df <- inner_join(df, genes_10k, by = "Trinity")
  
  # Assign the updated data frame back to its original name
  assign(name, df)
}

```




```{r}

# Add a day column to each filtered data frame
hpe12_filtered$Day <- "HPE12"
dpe1_filtered$Day <- "DPE1"
dpe3_filtered$Day <- "DPE3"
dpe7_filtered$Day <- "DPE7"
dpe14_filtered$Day <- "DPE14"
dpe14a_filtered$Day <- "DPE14a"
dpe14p_filtered$Day <- "DPE14p"
dpe21_filtered$Day <- "DPE21"

# Select the relevant columns
hpe12_filtered <- hpe12_filtered[, c("Trinity", "log2FoldChange", "padj", "Day")]
dpe1_filtered <- dpe1_filtered[, c("Trinity", "log2FoldChange", "padj", "Day")]
dpe3_filtered <- dpe3_filtered[, c("Trinity", "log2FoldChange", "padj", "Day")]
dpe7_filtered <- dpe7_filtered[, c("Trinity", "log2FoldChange", "padj", "Day")]
dpe14_filtered <- dpe14_filtered[, c("Trinity", "log2FoldChange", "padj", "Day")]
dpe14a_filtered <- dpe14a_filtered[, c("Trinity", "log2FoldChange", "padj", "Day")]
dpe14p_filtered <- dpe14p_filtered[, c("Trinity", "log2FoldChange", "padj", "Day")]
dpe21_filtered <- dpe21_filtered[, c("Trinity", "log2FoldChange", "padj", "Day")]



# Combine the data frames into one
combined_df <- rbind(hpe12_filtered, dpe1_filtered, dpe3_filtered, dpe7_filtered, 
                     dpe14_filtered, dpe14a_filtered, dpe14p_filtered, dpe21_filtered)

# Load ggplot2 for plotting
library(ggplot2)

# Create the jitter plot with labels for log2FoldChange > 10
ggplot(combined_df, aes(x = Day, y = log2FoldChange)) +
  geom_jitter(width = 0.2, height = 0) +
  geom_text(aes(label = ifelse(log2FoldChange > 7, Trinity, '')), 
            hjust = 0, vjust = 0, size = 3, check_overlap = TRUE) +
  labs(title = "Gene Expression Jitter Plot",
       x = "Day of Expression") +
  theme_minimal()

```


# 5K


```{r}

# Extract the IDs from the second column of the TSV file
trxpts_5k_ids <- genes_5k$Trinity

# Apply the filter function to each CSV data frame
hpe12_5k_filtered <- filter_csv(hpe12, trxpts_5k_ids)
dpe1_5k_filtered <- filter_csv(dpe1, trxpts_5k_ids)
dpe3_5k_filtered <- filter_csv(dpe3, trxpts_5k_ids)
dpe7_5k_filtered <- filter_csv(dpe7, trxpts_5k_ids)
dpe14_5k_filtered <- filter_csv(dpe14, trxpts_5k_ids)
dpe14a_5k_filtered <- filter_csv(dpe14a, trxpts_5k_ids)
dpe14p_5k_filtered <- filter_csv(dpe14p, trxpts_5k_ids)
dpe21_5k_filtered <- filter_csv(dpe21, trxpts_5k_ids)

#DROP NA ROWS

# Apply this function to each filtered data frame
hpe12_5k_filtered <- drop_na_rows(hpe12_5k_filtered)
dpe1_5k_filtered <- drop_na_rows(dpe1_5k_filtered)
dpe3_5k_filtered <- drop_na_rows(dpe3_5k_filtered)
dpe7_5k_filtered <- drop_na_rows(dpe7_5k_filtered)
dpe14_5k_filtered <- drop_na_rows(dpe14_5k_filtered)
dpe14a_5k_filtered <- drop_na_rows(dpe14a_5k_filtered)
dpe14p_5k_filtered <- drop_na_rows(dpe14p_5k_filtered)
dpe21_5k_filtered <- drop_na_rows(dpe21_5k_filtered)


```


# 2K


```{r}

# Extract the IDs from the second column of the TSV file
trxpts_2k_ids <- genes_2k$Trinity

# Apply the filter function to each CSV data frame
hpe12_2k_filtered <- filter_csv(hpe12, trxpts_2k_ids)
dpe1_2k_filtered <- filter_csv(dpe1, trxpts_2k_ids)
dpe3_2k_filtered <- filter_csv(dpe3, trxpts_2k_ids)
dpe7_2k_filtered <- filter_csv(dpe7, trxpts_2k_ids)
dpe14_2k_filtered <- filter_csv(dpe14, trxpts_2k_ids)
dpe14a_2k_filtered <- filter_csv(dpe14a, trxpts_2k_ids)
dpe14p_2k_filtered <- filter_csv(dpe14p, trxpts_2k_ids)
dpe21_2k_filtered <- filter_csv(dpe21, trxpts_2k_ids)

#DROP NA ROWS

# Apply this function to each filtered data frame
hpe12_2k_filtered <- drop_na_rows(hpe12_2k_filtered)
dpe1_2k_filtered <- drop_na_rows(dpe1_2k_filtered)
dpe3_2k_filtered <- drop_na_rows(dpe3_2k_filtered)
dpe7_2k_filtered <- drop_na_rows(dpe7_2k_filtered)
dpe14_2k_filtered <- drop_na_rows(dpe14_2k_filtered)
dpe14a_2k_filtered <- drop_na_rows(dpe14a_2k_filtered)
dpe14p_2k_filtered <- drop_na_rows(dpe14p_2k_filtered)
dpe21_2k_filtered <- drop_na_rows(dpe21_2k_filtered)


```


# ADJACENT


```{r}

# Extract the IDs from the second column of the TSV file
trxpts_adjacent_ids <- genes_adjacent$Trinity

# Apply the filter function to each CSV data frame
hpe12_adjacent_filtered <- filter_csv(hpe12, trxpts_adjacent_ids)
dpe1_adjacent_filtered <- filter_csv(dpe1, trxpts_adjacent_ids)
dpe3_adjacent_filtered <- filter_csv(dpe3, trxpts_adjacent_ids)
dpe7_adjacent_filtered <- filter_csv(dpe7, trxpts_adjacent_ids)
dpe14_adjacent_filtered <- filter_csv(dpe14, trxpts_adjacent_ids)
dpe14a_adjacent_filtered <- filter_csv(dpe14a, trxpts_adjacent_ids)
dpe14p_adjacent_filtered <- filter_csv(dpe14p, trxpts_adjacent_ids)
dpe21_adjacent_filtered <- filter_csv(dpe21, trxpts_adjacent_ids)

#DROP NA ROWS

# Apply this function to each filtered data frame
hpe12_adjacent_filtered <- drop_na_rows(hpe12_adjacent_filtered)
dpe1_adjacent_filtered <- drop_na_rows(dpe1_adjacent_filtered)
dpe3_adjacent_filtered <- drop_na_rows(dpe3_adjacent_filtered)
dpe7_adjacent_filtered <- drop_na_rows(dpe7_adjacent_filtered)
dpe14_adjacent_filtered <- drop_na_rows(dpe14_adjacent_filtered)
dpe14a_adjacent_filtered <- drop_na_rows(dpe14a_adjacent_filtered)
dpe14p_adjacent_filtered <- drop_na_rows(dpe14p_adjacent_filtered)
dpe21_adjacent_filtered <- drop_na_rows(dpe21_adjacent_filtered)


```


## MAPPING TO GENE SYMBOLS

```{r}

library(readr)
library(dplyr)

hsap_blastp.df <- read_delim(
  file = "/Users/joshuagmedina/Dropbox/01-DEVNEURO-LAB/34-scRNA-seq/08-SCRNA-HGLABv4/00_REFS/02_HUMAN_NCBI/augustus_hints_w_utr.v4.human_ncbi.map.tsv", 
  delim = "\t", 
  col_names = FALSE)

colnames(hsap_blastp.df) <- c("qseqid", "product_accession")


```


```{r}

human_ncbi.metadata <- read_delim(
  file = "/Users/joshuagmedina/Dropbox/02-PHD_THESIS/07-REF-GENOME-ARTICLE/01-ANALYSES/00-DATA/03-REFGENOME_V1_FUNCTIONAL_ANNOT/02-FUNCTIONAL_ANNOT_REF_HUMANS/00-REF/GCF_000001405.40_GRCh38.p14_feature_table.txt", 
  delim = "\t", 
  col_names = TRUE)

head(human_ncbi.metadata)

```

```{r}

human_ncbi.metadata_subset <- human_ncbi.metadata %>% filter(class == "with_protein") %>% dplyr::select(product_accession, symbol)

head(human_ncbi.metadata_subset)

```

```{r}

hsap_blastp.df <- hsap_blastp.df[!duplicated(hsap_blastp.df$qseqid), ]

hsap_mapping_df_tmp <- inner_join(hsap_blastp.df, human_ncbi.metadata_subset, by = "product_accession")
hasap_geneID_mapping_df <- hsap_mapping_df_tmp %>% dplyr::select(qseqid, symbol)

head(hasap_geneID_mapping_df)

```




```{r}


hsap_mapping_df_tmp <- inner_join(hsap_blastp.df, human_ncbi.metadata_subset, by = "product_accession")
colnames(hsap_mapping_df_tmp) <- c("genome_pred_id", "product_accession", "symbol")

```

 
# MAPPING TO DF DATA

## 10K

```{r}

# Define the function
merge_and_fill <- function(df_A, df_B) {
  result <- left_join(df_A, df_B, by = "genome_pred_id")
  
  result <- result %>%
    mutate(product_accession = ifelse(is.na(product_accession), genome_pred_id, product_accession),
           symbol = ifelse(is.na(symbol), genome_pred_id, symbol))
  
  return(result)
}

hpe12_filtered <- merge_and_fill(hpe12_filtered, hsap_mapping_df_tmp)
dpe1_filtered <- merge_and_fill(dpe1_filtered, hsap_mapping_df_tmp)
dpe3_filtered <- merge_and_fill(dpe3_filtered, hsap_mapping_df_tmp)
dpe7_filtered <- merge_and_fill(dpe7_filtered, hsap_mapping_df_tmp)
dpe14_filtered <- merge_and_fill(dpe14_filtered, hsap_mapping_df_tmp)
dpe14a_filtered <- merge_and_fill(dpe14a_filtered, hsap_mapping_df_tmp)
dpe14p_filtered <- merge_and_fill(dpe14p_filtered, hsap_mapping_df_tmp)
dpe21_filtered <- merge_and_fill(dpe21_filtered, hsap_mapping_df_tmp)



```


```{r}



# Add a day column to each filtered data frame
hpe12_filtered$Day <- "HPE12"
dpe1_filtered$Day <- "DPE1"
dpe3_filtered$Day <- "DPE3"
dpe7_filtered$Day <- "DPE7"
dpe14_filtered$Day <- "DPE14"
dpe14a_filtered$Day <- "DPE14a"
dpe14p_filtered$Day <- "DPE14p"
dpe21_filtered$Day <- "DPE21"

# Select the relevant columns
hpe12_filtered <- hpe12_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe1_filtered <- dpe1_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe3_filtered <- dpe3_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe7_filtered <- dpe7_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe14_filtered <- dpe14_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe14a_filtered <- dpe14a_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe14p_filtered <- dpe14p_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe21_filtered <- dpe21_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]

# Combine the data frames into one
dist10k <- rbind(hpe12_filtered, dpe1_filtered, dpe3_filtered, dpe7_filtered, 
                     dpe14_filtered, dpe14a_filtered, dpe14p_filtered, dpe21_filtered)


dist10k_filter <- dist10k[dist10k$log2FoldChange <= 12, ]

# Create the jitter plot with labels for log2FoldChange > 7
ggplot(dist10k_filter, aes(x = Day, y = log2FoldChange)) +
  geom_jitter(width = 0.2, height = 0) +
  geom_text_repel(aes(label = ifelse(log2FoldChange > 5, symbol, '')),
                  size = 3,
                  box.padding = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50') +
  labs(title = "Gene Expression Jitter Plot",
       x = "Day of Expression",
       y = "log2FoldChange") +
  theme_minimal()

```

## ADJACENT GENES

```{r}

# List of data frame names as strings
df_adj <- c(
  'hpe12_adjacent_filtered', 'dpe1_adjacent_filtered', 'dpe3_adjacent_filtered', 'dpe7_adjacent_filtered',
  'dpe14_adjacent_filtered', 'dpe14a_adjacent_filtered', 'dpe14p_adjacent_filtered', 'dpe21_adjacent_filtered')

# Loop through each data frame name and perform the join operation
for (name in df_adj) {
  # Get the data frame by name
  df <- get(name)
  
  # Perform the inner join with genes_10k
  df <- inner_join(df, genes_adjacent, by = "Trinity")
  
  # Assign the updated data frame back to its original name
  assign(name, df)
}


```



```{r}


hpe12_adjacent_filtered <- merge_and_fill(hpe12_adjacent_filtered, hsap_mapping_df_tmp)
dpe1_adjacent_filtered <- merge_and_fill(dpe1_adjacent_filtered, hsap_mapping_df_tmp)
dpe3_adjacent_filtered <- merge_and_fill(dpe3_adjacent_filtered, hsap_mapping_df_tmp)
dpe7_adjacent_filtered <- merge_and_fill(dpe7_adjacent_filtered, hsap_mapping_df_tmp)
dpe14_adjacent_filtered <- merge_and_fill(dpe14_adjacent_filtered, hsap_mapping_df_tmp)
dpe14a_adjacent_filtered <- merge_and_fill(dpe14a_adjacent_filtered, hsap_mapping_df_tmp)
dpe14p_adjacent_filtered <- merge_and_fill(dpe14p_adjacent_filtered, hsap_mapping_df_tmp)
dpe21_adjacent_filtered <- merge_and_fill(dpe21_adjacent_filtered, hsap_mapping_df_tmp)



```


```{r}



# Add a day column to each filtered data frame
hpe12_adjacent_filtered$Day <- "HPE12"
dpe1_adjacent_filtered$Day <- "DPE1"
dpe3_adjacent_filtered$Day <- "DPE3"
dpe7_adjacent_filtered$Day <- "DPE7"
dpe14_adjacent_filtered$Day <- "DPE14"
dpe14a_adjacent_filtered$Day <- "DPE14a"
dpe14p_adjacent_filtered$Day <- "DPE14p"
dpe21_adjacent_filtered$Day <- "DPE21"

# Select the relevant columns
hpe12_adjacent_filtered <- hpe12_adjacent_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe1_adjacent_filtered <- dpe1_adjacent_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe3_adjacent_filtered <- dpe3_adjacent_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe7_adjacent_filtered <- dpe7_adjacent_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe14_adjacent_filtered <- dpe14_adjacent_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe14a_adjacent_filtered <- dpe14a_adjacent_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe14p_adjacent_filtered <- dpe14p_adjacent_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]
dpe21_adjacent_filtered <- dpe21_adjacent_filtered[, c("Trinity", "log2FoldChange", "padj", "Day","genome_pred_id", "product_accession", "symbol")]

# Combine the data frames into one
dist_adj <- rbind(hpe12_adjacent_filtered, dpe1_adjacent_filtered, dpe3_adjacent_filtered, dpe7_adjacent_filtered, 
                     dpe14_adjacent_filtered, dpe14a_adjacent_filtered, dpe14p_adjacent_filtered, dpe21_adjacent_filtered)


dist_adj_filter <- dist_adj[dist_adj$log2FoldChange <= 12, ]

# Define the order of the levels for the Day variable
reg_order <- c("HPE12", "DPE1", "DPE3", "DPE7", "DPE14", "DPE14a", "DPE14p", "DPE21")

# Convert the Day variable to a factor with the specified levels
dist_adj_filter$Day <- factor(dist_adj_filter$Day, levels = reg_order)

# Create the jitter plot with labels for log2FoldChange > 7
ggplot(dist_adj_filter, aes(x = Day, y = log2FoldChange)) +
  geom_jitter(width = 0.2, height = 0) +
  geom_text_repel(aes(label = ifelse(log2FoldChange > 5, symbol, '')),
                  size = 3,
                  box.padding = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50') +
  labs(title = "Gene Expression Jitter Plot",
       x = "Day of Expression",
       y = "log2FoldChange") +
  theme_minimal()

```




## Percentage of total genes identified

```{r}

# List of data frame names as strings
df_10k <- c(
  'hpe12_filtered', 'dpe1_filtered', 'dpe3_filtered', 'dpe7_filtered',
  'dpe14_filtered', 'dpe14a_filtered', 'dpe14p_filtered', 'dpe21_filtered')

# Create a list of data frames
data_10k <- mget(df_10k)

# Count the number of observations in each data frame
no_expressed_10k <- sapply(data_10k, nrow)

# Create a summary data frame
data10k_summary <- data.frame(
  DataFrame = names(no_expressed_10k),
  TotalObservations = no_expressed_10k
)

data10k_summary <- data10k_summary %>%
  mutate(Percentage_of_Total = (TotalObservations / 14955) * 100)


# Recode the values in the DataFrame column
data10k_summary <- data10k_summary %>%
  mutate(DataFrame = recode(DataFrame,
                            'hpe12_filtered' = '12HPE',
                            'dpe1_filtered' = '1DPE',
                            'dpe3_filtered' = '3DPE',
                            'dpe7_filtered' = '7DPE',
                            'dpe14_filtered' = '14DPE',
                            'dpe14a_filtered' = '14aDPE',
                            'dpe14p_filtered' = '14pDPE',
                            'dpe21_filtered' = '21DPE'))

colnames(data10k_summary) <- c("day_of_regeneration", "upregulated_genes", "percentage_of_total")


# Print the summary data frame
print(data10k_summary)

```

```{r}

# List of data frame names as strings
df_5k <- c(
  'hpe12_5k_filtered', 'dpe1_5k_filtered', 'dpe3_5k_filtered', 'dpe7_5k_filtered',
  'dpe14_5k_filtered', 'dpe14a_5k_filtered', 'dpe14p_5k_filtered', 'dpe21_5k_filtered')

# Create a list of data frames
data_5k <- mget(df_5k)

# Count the number of observations in each data frame
no_expressed_5k <- sapply(data_5k, nrow)

# Create a summary data frame
data5k_summary <- data.frame(
  DataFrame = names(no_expressed_5k),
  TotalObservations = no_expressed_5k
)

data5k_summary <- data5k_summary %>%
  mutate(Percentage_of_Total = (TotalObservations / 12180) * 100)


# Recode the values in the DataFrame column
data5k_summary <- data5k_summary %>%
  mutate(DataFrame = recode(DataFrame,
                            'hpe12_5k_filtered' = '12HPE',
                            'dpe1_5k_filtered' = '1DPE',
                            'dpe3_5k_filtered' = '3DPE',
                            'dpe7_5k_filtered' = '7DPE',
                            'dpe14_5k_filtered' = '14DPE',
                            'dpe14a_5k_filtered' = '14aDPE',
                            'dpe14p_5k_filtered' = '14pDPE',
                            'dpe21_5k_filtered' = '21DPE'))

colnames(data5k_summary) <- c("day_of_regeneration", "upregulated_genes", "percentage_of_total")


# Print the summary data frame
print(data5k_summary)

```



```{r}

# List of data frame names as strings
df_2k <- c(
  'hpe12_2k_filtered', 'dpe1_2k_filtered', 'dpe3_2k_filtered', 'dpe7_2k_filtered',
  'dpe14_2k_filtered', 'dpe14a_2k_filtered', 'dpe14p_2k_filtered', 'dpe21_2k_filtered')

# Create a list of data frames
data_2k <- mget(df_2k)

# Count the number of observations in each data frame
no_expressed_2k <- sapply(data_2k, nrow)

# Create a summary data frame
data2k_summary <- data.frame(
  DataFrame = names(no_expressed_2k),
  TotalObservations = no_expressed_2k
)

data2k_summary <- data2k_summary %>%
  mutate(Percentage_of_Total = (TotalObservations / 10312) * 100)


# Recode the values in the DataFrame column
data2k_summary <- data2k_summary %>%
  mutate(DataFrame = recode(DataFrame,
                            'hpe12_2k_filtered' = '12HPE',
                            'dpe1_2k_filtered' = '1DPE',
                            'dpe3_2k_filtered' = '3DPE',
                            'dpe7_2k_filtered' = '7DPE',
                            'dpe14_2k_filtered' = '14DPE',
                            'dpe14a_2k_filtered' = '14aDPE',
                            'dpe14p_2k_filtered' = '14pDPE',
                            'dpe21_2k_filtered' = '21DPE'))

colnames(data2k_summary) <- c("day_of_regeneration", "upregulated_genes", "percentage_of_total")


# Print the summary data frame
print(data2k_summary)

```



```{r}

# List of data frame names as strings
df_adj <- c(
  'hpe12_adjacent_filtered', 'dpe1_adjacent_filtered', 'dpe3_adjacent_filtered', 'dpe7_adjacent_filtered',
  'dpe14_adjacent_filtered', 'dpe14a_adjacent_filtered', 'dpe14p_adjacent_filtered', 'dpe21_adjacent_filtered')

# Create a list of data frames
data_adj <- mget(df_adj)

# Count the number of observations in each data frame
no_expressed_adj <- sapply(data_adj, nrow)

# Create a summary data frame
data_adj_summary <- data.frame(
  DataFrame = names(no_expressed_adj),
  TotalObservations = no_expressed_adj
)

data_adj_summary <- data_adj_summary %>%
  mutate(Percentage_of_Total = (TotalObservations / 12498) * 100)


# Recode the values in the DataFrame column
data_adj_summary <- data_adj_summary %>%
  mutate(DataFrame = recode(DataFrame,
                            'hpe12_adjacent_filtered' = '12HPE',
                            'dpe1_adjacent_filtered' = '1DPE',
                            'dpe3_adjacent_filtered' = '3DPE',
                            'dpe7_adjacent_filtered' = '7DPE',
                            'dpe14_adjacent_filtered' = '14DPE',
                            'dpe14a_adjacent_filtered' = '14aDPE',
                            'dpe14p_adjacent_filtered' = '14pDPE',
                            'dpe21_adjacent_filtered' = '21DPE'))

colnames(data_adj_summary) <- c("day_of_regeneration", "upregulated_genes", "percentage_of_total")


# Print the summary data frame
print(data_adj_summary)

```



## Percentage of upregulated by distance

```{r}

library(RColorBrewer)

# Add a column to indicate the source of each data frame
data10k_summary <- data10k_summary %>% mutate(Distance = '10Kb')
data5k_summary <- data5k_summary %>% mutate(Distance = '5Kb')
data2k_summary <- data2k_summary %>% mutate(Distance = '2Kb')
data_adj_summary <- data_adj_summary %>% mutate(Distance = 'Adjacent')

# Combine all data frames
combined_summary <- bind_rows(data10k_summary, data5k_summary, data2k_summary, data_adj_summary)

# Ensure the 'Day' column is treated as a factor with the correct order if necessary
combined_summary$day_of_regeneration <- factor(combined_summary$day_of_regeneration, levels = c("12HPE","1DPE", "3DPE", "7DPE", "14DPE", "14aDPE", "14pDPE", "21DPE"))
combined_summary$Distance <- factor(combined_summary$Distance, levels = c("2Kb","5Kb", "3DPE", "10Kb", "Adjacent"))

# Create the bar plot
genes_expressed_per_distance <- ggplot(combined_summary, aes(x = day_of_regeneration, y = percentage_of_total, fill = Distance)) +
  geom_bar(stat = 'identity', position = 'dodge') + scale_fill_brewer(palette = "Paired") + 
  labs(x = "Day of Regeneration",
       y = "Percentage of Total",
       fill = "Distance") +
  theme_classic() +
  theme(axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 16, color = "black", face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.key.size = unit(1.5, "lines")) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1))
genes_expressed_per_distance

```


## Percentage of total genes identified

```{r}

# List of data frame names as strings
df_10k <- c(
  'hpe12_filtered', 'dpe1_filtered', 'dpe3_filtered', 'dpe7_filtered',
  'dpe14_filtered', 'dpe14a_filtered', 'dpe14p_filtered', 'dpe21_filtered')

# Create a list of data frames
data_10k <- mget(df_10k)

# Count the number of observations in each data frame
no_expressed_10k <- sapply(data_10k, function(df) n_distinct(df$Trinity))

# Create a summary data frame
data10k_summary <- data.frame(
  DataFrame = names(no_expressed_10k),
  TotalObservations = no_expressed_10k
)

data10k_summary <- data10k_summary %>%
  mutate(Percentage_of_Total = (TotalObservations / 14955) * 100)


# Recode the values in the DataFrame column
data10k_summary <- data10k_summary %>%
  mutate(DataFrame = recode(DataFrame,
                            'hpe12_filtered' = '12HPE',
                            'dpe1_filtered' = '1DPE',
                            'dpe3_filtered' = '3DPE',
                            'dpe7_filtered' = '7DPE',
                            'dpe14_filtered' = '14DPE',
                            'dpe14a_filtered' = '14aDPE',
                            'dpe14p_filtered' = '14pDPE',
                            'dpe21_filtered' = '21DPE'))

colnames(data10k_summary) <- c("day_of_regeneration", "upregulated_genes", "percentage_of_total")


# Print the summary data frame
print(data10k_summary)

```

```{r}

# List of data frame names as strings
df_5k <- c(
  'hpe12_5k_filtered', 'dpe1_5k_filtered', 'dpe3_5k_filtered', 'dpe7_5k_filtered',
  'dpe14_5k_filtered', 'dpe14a_5k_filtered', 'dpe14p_5k_filtered', 'dpe21_5k_filtered')

# Create a list of data frames
data_5k <- mget(df_5k)

# Count the number of observations in each data frame
no_expressed_5k <- sapply(data_5k, function(df) n_distinct(df$Trinity))

# Create a summary data frame
data5k_summary <- data.frame(
  DataFrame = names(no_expressed_5k),
  TotalObservations = no_expressed_5k
)

data5k_summary <- data5k_summary %>%
  mutate(Percentage_of_Total = (TotalObservations / 12180) * 100)


# Recode the values in the DataFrame column
data5k_summary <- data5k_summary %>%
  mutate(DataFrame = recode(DataFrame,
                            'hpe12_5k_filtered' = '12HPE',
                            'dpe1_5k_filtered' = '1DPE',
                            'dpe3_5k_filtered' = '3DPE',
                            'dpe7_5k_filtered' = '7DPE',
                            'dpe14_5k_filtered' = '14DPE',
                            'dpe14a_5k_filtered' = '14aDPE',
                            'dpe14p_5k_filtered' = '14pDPE',
                            'dpe21_5k_filtered' = '21DPE'))

colnames(data5k_summary) <- c("day_of_regeneration", "upregulated_genes", "percentage_of_total")


# Print the summary data frame
print(data5k_summary)

```



```{r}

# List of data frame names as strings
df_2k <- c(
  'hpe12_2k_filtered', 'dpe1_2k_filtered', 'dpe3_2k_filtered', 'dpe7_2k_filtered',
  'dpe14_2k_filtered', 'dpe14a_2k_filtered', 'dpe14p_2k_filtered', 'dpe21_2k_filtered')

# Create a list of data frames
data_2k <- mget(df_2k)

# Count the number of observations in each data frame
no_expressed_2k <- sapply(data_2k, function(df) n_distinct(df$Trinity))

# Create a summary data frame
data2k_summary <- data.frame(
  DataFrame = names(no_expressed_2k),
  TotalObservations = no_expressed_2k
)

data2k_summary <- data2k_summary %>%
  mutate(Percentage_of_Total = (TotalObservations / 10312) * 100)


# Recode the values in the DataFrame column
data2k_summary <- data2k_summary %>%
  mutate(DataFrame = recode(DataFrame,
                            'hpe12_2k_filtered' = '12HPE',
                            'dpe1_2k_filtered' = '1DPE',
                            'dpe3_2k_filtered' = '3DPE',
                            'dpe7_2k_filtered' = '7DPE',
                            'dpe14_2k_filtered' = '14DPE',
                            'dpe14a_2k_filtered' = '14aDPE',
                            'dpe14p_2k_filtered' = '14pDPE',
                            'dpe21_2k_filtered' = '21DPE'))

colnames(data2k_summary) <- c("day_of_regeneration", "upregulated_genes", "percentage_of_total")


# Print the summary data frame
print(data2k_summary)

```



```{r}

# List of data frame names as strings
df_adj <- c(
  'hpe12_adjacent_filtered', 'dpe1_adjacent_filtered', 'dpe3_adjacent_filtered', 'dpe7_adjacent_filtered',
  'dpe14_adjacent_filtered', 'dpe14a_adjacent_filtered', 'dpe14p_adjacent_filtered', 'dpe21_adjacent_filtered')

# Create a list of data frames
data_adj <- mget(df_adj)

# Count the number of observations in each data frame
no_expressed_adj <- sapply(data_adj, function(df) n_distinct(df$Trinity))

# Create a summary data frame
data_adj_summary <- data.frame(
  DataFrame = names(no_expressed_adj),
  TotalObservations = no_expressed_adj
)

data_adj_summary <- data_adj_summary %>%
  mutate(Percentage_of_Total = (TotalObservations / 12498) * 100)


# Recode the values in the DataFrame column
data_adj_summary <- data_adj_summary %>%
  mutate(DataFrame = recode(DataFrame,
                            'hpe12_adjacent_filtered' = '12HPE',
                            'dpe1_adjacent_filtered' = '1DPE',
                            'dpe3_adjacent_filtered' = '3DPE',
                            'dpe7_adjacent_filtered' = '7DPE',
                            'dpe14_adjacent_filtered' = '14DPE',
                            'dpe14a_adjacent_filtered' = '14aDPE',
                            'dpe14p_adjacent_filtered' = '14pDPE',
                            'dpe21_adjacent_filtered' = '21DPE'))

colnames(data_adj_summary) <- c("day_of_regeneration", "upregulated_genes", "percentage_of_total")


# Print the summary data frame
print(data_adj_summary)

```



## Percentage of upregulated by distance

```{r}

library(RColorBrewer)

# Add a column to indicate the source of each data frame
data10k_summary <- data10k_summary %>% mutate(Distance = '10Kb')
data5k_summary <- data5k_summary %>% mutate(Distance = '5Kb')
data2k_summary <- data2k_summary %>% mutate(Distance = '2Kb')
data_adj_summary <- data_adj_summary %>% mutate(Distance = 'Adjacent')

# Combine all data frames
combined_summary <- bind_rows(data10k_summary, data5k_summary, data2k_summary, data_adj_summary)

# Ensure the 'Day' column is treated as a factor with the correct order if necessary
combined_summary$day_of_regeneration <- factor(combined_summary$day_of_regeneration, levels = c("12HPE","1DPE", "3DPE", "7DPE", "14DPE", "14aDPE", "14pDPE", "21DPE"))
combined_summary$Distance <- factor(combined_summary$Distance, levels = c("2Kb","5Kb", "Adjacent", "10Kb"))

# Create the bar plot
genes_expressed_per_distance <- ggplot(combined_summary, aes(x = day_of_regeneration, y = percentage_of_total, fill = Distance)) +
  geom_bar(stat = 'identity', position = 'dodge') + scale_fill_brewer(palette = "Paired") + 
  labs(x = "Day of Regeneration",
       y = "% of Upregulated Genes",
       fill = "Distance") +
  theme_classic() +
  theme(axis.text = element_text(size = 10, color = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_text(size = 16, color = "black", face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.key.size = unit(1.5, "lines")) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1))
genes_expressed_per_distance

```



```{r}

library(RColorBrewer)

# Add a column to indicate the source of each data frame
data10k_summary <- data10k_summary %>% mutate(Distance = '10Kb')
data5k_summary <- data5k_summary %>% mutate(Distance = '5Kb')
data2k_summary <- data2k_summary %>% mutate(Distance = '2Kb')

# Combine all data frames
combined_summary <- bind_rows(data10k_summary, data5k_summary, data2k_summary)

combined_summary <- combined_summary %>%
  filter(day_of_regeneration %in% c("12HPE", "1DPE", "3DPE", "7DPE", "14DPE", "14aDPE", "14pDPE", "21DPE"))

# Ensure the 'Day' column is treated as a factor with the correct order if necessary
combined_summary$day_of_regeneration <- factor(combined_summary$day_of_regeneration, levels = c("12HPE","1DPE", "3DPE", "7DPE", "14DPE", "14aDPE", "14pDPE", "21DPE"))
combined_summary$Distance <- factor(combined_summary$Distance, levels = c("2Kb","5Kb", "10Kb"))

# Create the bar plot
genes_expressed_per_distance <- ggplot(combined_summary, aes(x = day_of_regeneration, y = percentage_of_total, fill = Distance)) +
  geom_bar(stat = 'identity', position = 'dodge') + scale_fill_brewer(palette = "Paired") + 
  labs(x = "Day of Regeneration",
       y = "% of Upregulated Genes",
       fill = "Distance") +
  theme_classic() +
  theme(axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 16, color = "black", face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.key.size = unit(1.5, "lines")) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1))
genes_expressed_per_distance

```






## Get expression of gypsy elements:

```{r}

library(stringr)

# Read the TSV file into a data frame
gypsy_ids <- read.delim("mashanov_gypsy_int_trxpt.tsv", header = TRUE)

# Extract the Trinity column
gypsy_trinity_strings <- gypsy_ids$Trinity
gypsy_values <- gypsy_ids$Gypsy

# List of data frames to check
data_frames <- list(hpe12 = hpe12, dpe1 = dpe1, dpe3 = dpe3, dpe7 = dpe7, dpe14 = dpe14, dpe14a = dpe14a, dpe14p = dpe14p, dpe21 = dpe21)

# Function to filter data frames based on exact substring match in Trinity column and add Gypsy values
filter_by_trinity <- function(df, gypsy_trinity_strings, gypsy_values) {
  # Initialize an empty data frame to store matched rows
  matched_df <- data.frame()
  
  # Iterate over each trinity string
  for (i in seq_along(gypsy_trinity_strings)) {
    # Create a pattern with word boundaries
    pattern <- paste0(gypsy_trinity_strings[i])
    
    # Debugging: Print the pattern to check if it's correct
    cat("Pattern:", pattern, "\n")
    
    # Find rows in the data frame where the Trinity column matches the pattern
    matches <- df %>% filter(str_detect(Trinity, pattern))
    
    # Debugging: Print the number of matches found
    cat("Number of matches for", gypsy_trinity_strings[i], ":", nrow(matches), "\n")
    
    # Add the corresponding Gypsy value to the matches
    if (nrow(matches) > 0) {
      matches$Gypsy <- gypsy_values[i]
      matched_df <- bind_rows(matched_df, matches)
    }
  }
  
  return(matched_df)
}

# Apply the function to each data frame and store results in a list
filtered_data_frames <- lapply(data_frames, filter_by_trinity, gypsy_trinity_strings = gypsy_trinity_strings, gypsy_values = gypsy_values)

```

## Get the most significant for each


```{r}

# Function to sort by Gypsy and padj, and keep only the row with the lowest padj for each Gypsy
sort_and_filter <- function(df) {
  df %>%
    arrange(Gypsy, padj) %>%
    group_by(Gypsy) %>%
    slice(1) %>%
    ungroup()
}

# Apply the sort and filter function to each filtered data frame
final_filtered_data_frames <- lapply(filtered_data_frames, sort_and_filter)


```

## Make a plot

```{r}

# Combine the data frames into a single data frame and add the timepoint column
combined_df <- bind_rows(
  lapply(names(final_filtered_data_frames), function(name) {
    df <- final_filtered_data_frames[[name]]
    df$timepoint <- strsplit(name, "_")[[1]][1]
    return(df)
  })
)


combined_df$timepoint <- factor(combined_df$timepoint, levels = unique(combined_df$timepoint))

# Extract the numeric part of the Gypsy IDs and use it to order the levels
combined_df <- combined_df %>%
  mutate(Gypsy_num = as.numeric(str_extract(Gypsy, "\\d+"))) %>%
  arrange(Gypsy_num) %>%
  mutate(Gypsy = factor(Gypsy, levels = unique(Gypsy)))

# Remove the temporary numeric column
combined_df <- combined_df %>% select(-Gypsy_num)

```

```{r}

# Generate the bar plot using ggplot2
ggplot(combined_df, aes(x = Gypsy, y = log2FoldChange, fill = Gypsy)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ timepoint, scales = "free_x") +
  labs(title = "Expression of Each Distinct Gypsy Over Regeneration",
       x = "Gypsy",
       y = "Expression") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  # Rotate x-axis labels for better readability


```


```{r}

ggplot(combined_df, aes(x = timepoint, y = Gypsy, fill = log2FoldChange)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = "",
       x = "Time Post Evisceration",
       y = "Gypsy",
       fill = "Expression") +
  theme_classic() +
  theme(axis.text.x = element_text(vjust = 0.5),
        axis.line = element_blank(),
        axis.ticks = element_blank())


```

```{r}

# Generate the volcano plot using ggplot2
ggplot(combined_df, aes(x = log2FoldChange, y = -log10(padj), label = Gypsy)) +
  geom_point(aes(color = timepoint), alpha = 0.6, size = 2) +
  geom_text_repel(size = 3, max.overlaps = 10) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Volcano Plot of Gypsy Expression",
       x = "Log2 Fold Change",
       y = "-Log10 Adjusted P-value",
       color = "Time of Regeneration") +
  theme_minimal()


combined_df <- combined_df %>%
  mutate(label_color = ifelse(abs(log2FoldChange) <= 2 & (padj >= 0.01 | is.na(padj)), "black", as.character(timepoint)))


# Generate the volcano plot using ggplot2
ggplot(combined_df, aes(x = log2FoldChange, y = -log10(padj), label = Gypsy)) +
  geom_point(aes(color = label_color), alpha = 0.6, size = 2) +
  scale_color_manual(values = c("black", RColorBrewer::brewer.pal(n = length(unique(combined_df$timepoint)), name = "Dark2"))) +
  geom_text_repel(data = combined_df %>% filter(label_color != "black"), 
                  aes(label = Gypsy, color = "black"), size = 3, max.overlaps = 10) +
  geom_text_repel(data = combined_df %>% filter(label_color == "black"), 
                  aes(label = Gypsy), color = "black", size = 3, max.overlaps = 10) +
  geom_vline(xintercept = c(-2, 2), linetype = "dashed", color = "grey40") +
  geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = "grey40") +
  labs(title = "",
       x = "Log2 Fold Change",
       y = "-Log10 Adjusted P-value",
       color = "Day of Regeneration") +
  theme_classic() +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title = element_text(size = 16, color = "black", face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.key.size = unit(1.5, "lines")) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1))



volcano_plot_gypsy <- ggplot(combined_df, aes(x = log2FoldChange, y = -log10(padj), label = Gypsy)) +
  geom_point(aes(color = label_color), alpha = 0.6, size = 2) +
  scale_color_manual(
    values = c("black", RColorBrewer::brewer.pal(n = length(unique(combined_df$timepoint)), name = "Dark2")),
    labels = c("not significant", levels(combined_df$timepoint))
  ) +
  geom_text_repel(data = combined_df %>% filter(label_color != "not significant"), 
                  aes(label = Gypsy), color = "black", size = 3, max.overlaps = 10) +
  geom_text_repel(data = combined_df %>% filter(label_color == "not significant"), 
                  aes(label = Gypsy), color = "black", size = 3, max.overlaps = 10) +
  geom_vline(xintercept = c(-2, 2), linetype = "dashed", color = "grey40") +
  geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = "grey40") +
  labs(title = "",
       x = "Log2 Fold Change",
       y = "-Log10 Adjusted P-value",
       color = "Day of Regeneration") +
  theme_classic() +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title = element_text(size = 16, color = "black", face = "bold"),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12, face = "bold"),
        legend.key.size = unit(1, "lines")) +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1))
volcano_plot_gypsy

```


```{r}

gypsy_expression <- volcano_plot_gypsy + genes_expressed_per_distance
gypsy_expression <- gypsy_expression + plot_annotation(tag_levels = "A")
gypsy_expression

```

# Histogram based on distance of adjacent genes.


```{r}

adjacent_genes_w_distance <- read.delim("gypsy_edta_closest_genes.tsv", header = FALSE)
colnames(adjacent_genes_w_distance) <- c("gypsy_scaffold", "gypsy_start","gypsy_end", "gypsy", "empty1", "gypsy_strand", "gene_scaffold", "gene_start", "gene_end", "genome_pred_id", "empty2", "strand", "gene_distance")


adjacent_genes_w_distance <- adjacent_genes_w_distance %>%
  select(genome_pred_id, gene_distance)

```

## DE Data Simplified

```{r}

# Create a function to extract specific columns
extract_columns <- function(df, columns) {
  return(df[, columns, drop = FALSE])
}

# List of data frames (replace with your actual data frames)
adjacent_expression_df <- list(hpe12_adjacent_filtered, dpe1_adjacent_filtered, dpe3_adjacent_filtered, 
                    dpe7_adjacent_filtered, dpe14_adjacent_filtered, dpe14a_adjacent_filtered,
                    dpe14p_adjacent_filtered, dpe21_adjacent_filtered)

# Define the columns to extract
columns_to_extract <- c("Trinity", "padj", "log2FoldChange")

# Use lapply to apply the function to each data frame
extracted_data_frames <- lapply(adjacent_expression_df, extract_columns, columns = columns_to_extract)

# Optional: Name the list elements for easier reference
names(extracted_data_frames) <- c("hpe12_adjacent_exp", "dpe1_adjacent_exp", "dpe3_adjacent_exp",
                                  "dpe7_adjacent_exp", "dpe14_adjacent_exp", "dpe14a_adjacent_exp",
                                  "dpe14p_adjacent_exp", "dpe21_adjacent_exp") # Adjust names accordingly

# Print the extracted data frames
print(extracted_data_frames)


```


```{r}

# Load necessary library
#library(dplyr)

# Define a function to join each CSV data frame with partial matching
join_df <- function(df, ids_df) {
  # Extract the part of the Trinity ID without the _i part
  df <- df %>%
    mutate(Trinity_short = sub("_i.*", "", Trinity))
  
  ids_df <- ids_df %>%
    mutate(Trinity_short = sub("_i.*", "", Trinity))
  
  # Perform the join with partial matching
  df_joined <- df %>%
    inner_join(ids_df, by = "Trinity_short")
  
  # Filter the resulting data frame
  df_filtered <- df_joined %>%
    filter(padj < 0.05 & log2FoldChange >= 2)
  
  # Remove the temporary Trinity_short column
  df_filtered <- df_filtered %>%
    select(-Trinity_short)
  
  return(df_filtered)
}


# Apply the function to each data frame in the list
joined_data_frames <- lapply(extracted_data_frames, join_df, ids_df = genes_adjacent)

# Print the joined and filtered data frames
head(joined_data_frames)


```

```{r}

# Load necessary library
#library(dplyr)

# Define a function to join each CSV data frame with partial matching
join_csv <- function(df, ids_df) {
  # Extract the part of the Trinity ID without the _i part
  df <- df %>%
    mutate(Trinity_short = sub("_i.*", "", Trinity))
  
  ids_df <- ids_df %>%
    mutate(Trinity_short = sub("_i.*", "", Trinity))
  
  # Perform the join with partial matching
  df_joined <- df %>%
    inner_join(ids_df, by = "Trinity_short")
  
  # Filter the resulting data frame
  df_filtered <- df_joined %>%
    filter(padj < 0.05 & log2FoldChange >= 2)
  
  # Remove the temporary Trinity_short column
  df_filtered <- df_filtered %>%
    select(-Trinity_short)
  
  return(df_filtered)
}

# Apply the function to each data frame in the list
joined_data_frames <- lapply(extracted_data_frames, join_csv, ids_df = genes_adjacent)

# Print the joined and filtered data frames
print(joined_data_frames)


```


```{r}

# Load necessary library
#library(dplyr)

# Define a function to join each data frame with gene data frame
join_with_gene_data <- function(df, gene_df) {
  # Ensure gene_df has unique genes
  gene_df <- gene_df %>%
    distinct(genome_pred_id, .keep_all = TRUE)
  
  # Perform the join
  df_joined <- df %>%
    inner_join(gene_df, by = c("genome_pred_id"))
  
  # Remove duplicates based on genome_pred_id column
  df_joined <- df_joined %>%
    distinct(genome_pred_id, .keep_all = TRUE)
  
  return(df_joined)
}


# Apply the function to each data frame in the list
final_adjacent_data_frames <- lapply(joined_data_frames, join_with_gene_data, gene_df = adjacent_genes_w_distance)

# Print the final data frames
head(final_adjacent_data_frames)


```


```{r}

for (i in seq_along(final_adjacent_data_frames)) {
  df_name <- names(final_adjacent_data_frames)[i]
  identifier <- sub("_.*", "", df_name)
  final_adjacent_data_frames[[i]] <- final_adjacent_data_frames[[i]] %>%
    mutate(source_df = identifier)
}

combined_adjacent_df <- bind_rows(final_adjacent_data_frames)

```


```{r}

# Define window size
window_size <- 2000

# Create distance intervals (bins)
combined_adjacent_df <- combined_adjacent_df %>%
  mutate(distance_bin = floor(gene_distance / window_size) * window_size)

# Calculate density of genes in each distance interval for each data frame
density_data <- combined_adjacent_df %>%
  group_by(source_df, distance_bin) %>%
  summarise(gene_count = n(), .groups = 'drop')


```


```{r}

# Plot the density using ggplot2 as a bar plot
gene_density_plot <- ggplot(density_data, aes(x = distance_bin, y = gene_count, fill = source_df)) +
  geom_bar(stat = "identity", position = "stack") + scale_fill_brewer(palette = "Dark2") +
  labs(
       x = "Distance (bin)",
       y = "Density of Genes",
       fill = "Day of Regeneration") +
  theme_bw() +
  theme(axis.text = element_text(size = 14, color = "black"),
        axis.title = element_text(size = 16, color = "black", face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16, face = "bold"),
        legend.key.size = unit(1.5, "lines"),
        legend.position = "bottom",
        legend.direction = "horizontal")
  

```

```{r}

# Plot the density using ggplot2 as a faceted bar plot
ggplot(density_data, aes(x = distance_bin, y = gene_count, fill = source_df)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ source_df, scales = "free_y") +
  labs(title = "Gene Density Plot by Source Data Frame",
       x = "Distance (bin)",
       y = "Density of Genes") +
  theme_minimal()


```

```{r}

# Combine the top two plots side by side
top_plots <- volcano_plot_gypsy + genes_expressed_per_distance + plot_layout(ncol = 2)

# Combine the top plots with the bottom plot
combined_plot <- top_plots / gene_density_plot
combined_plot <- combined_plot + plot_annotation(tag_levels = "A")
# Print the combined plot
print(combined_plot)

```

